Partial reproduction of Bush et al (2017) Oxic-anoxic regime shifts mediated by feedbacks between biogeochemical processes and microbial community dynamics

### 2021-04-14 various_stuff/dev_vignettes/partial reproduction Bushetal2017.Rmd
reorganisation of vignettes
now need to get them working!

---
title: "Diversity and the nature of responses to environmental change."
author: "Owen Petchey & Jennifer Probst"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 4
    number_sections: true
    code_folding: hide
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Setup R

@@

library(tidyverse)
library(deSolve)
library(rootSolve)

set.seed(13)

run_sims <- FALSE
```

@@

@@

  ## oxygen production per microbial cell
  P_CB = 6e-9,

@@

@@

Used to add noise and to ensure minimum abundance. Noise is added to the resource concentrations, and not to the organism abundances. Minimum organism abundance is set. I.e. organismal abundance cannot go below some very low level.

```{r}
source("functions/default_event_definition.r")
```

@@

This is when an event happens

@@

Default to no noise:

@@

Default to following minimums

@@

Here we specify no forcing of a_0 (used unless set elsewhere).

Initial conditions favouring anoxic.

## Simulation function

```{r}
source("functions/run_simulation.r")
```

## Helper functions

```{r}
source("functions/plot_dynamics.r")
source("functions/ss_expt_a_N.r")
source("functions/get_nonlinearity.r")
source("functions/get_hysteresis_range.r")
```

@@

```{r}
# dynamic_model = bushplus_dynamic_model
# event_definition = default_event_definition
# parameter_values = default_parameter_values
# event_interval = default_event_interval
# noise_sigma = default_noise_sigma
# minimum_abundances = default_minimum_abundances
# sim_duration = default_sim_duration
# sim_sample_interval = default_sim_sample_interval
# log10a_forcing = default_log10a_forcing
# initial_state = default_initial_state
# solver_method = "radau"
# 
# 
# times <- seq(0,
#              sim_duration,
#              by = sim_sample_interval)
# 
# event_times <- seq(min(times),
#                    max(times),
#                    by = event_interval)  
# 
# log10a_forcing <- approxfun(x = log10a_forcing[,1],
#                             y = log10a_forcing[,2],
#                             method = "linear",
#                             rule = 2)
# assign("log10a_forcing", log10a_forcing, envir = .GlobalEnv)
# 
# out <- as.data.frame(ode(y = initial_state,
#                          times = times,
#                          func = dynamic_model,
#                          parms = parameter_values,
#                          method = solver_method,
#                          events = list(func=event_definition,
#                                        time=event_times)))
# 
# sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
#                            event_definition = default_event_definition,
#                            parameter_values = default_parameter_values,
#                            event_interval = default_event_interval,
#                            noise_sigma = default_noise_sigma,
#                            minimum_abundances = default_minimum_abundances,
#                            sim_duration = default_sim_duration,
#                            sim_sample_interval = default_sim_sample_interval,
#                            log10a_forcing = default_log10a_forcing,
#                            initial_state = default_initial_state,
#                            solver_method = "radau")

#sim_res <- run_simulation()

#plot_dynamics(sim_res)

@@

# Results: Reproducing previous results

@@

sim_res <- run_simulation(initial_state = init_state_favouring_anoxic)
plot_dynamics(simulation_result = sim_res)

@@

## Initial conditions favouring anoxic, with noise, no forcing

Figure 2a, b in Bush et al, but with noise

```{r}
## add some noise
noise_sigma_0.01 <- 0.01
sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
                          noise_sigma = noise_sigma_0.01)
plot_dynamics(sim_res)
#ggsave("high_SB_PB.png")
```

@@

sim_res <- run_simulation(initial_state = init_state_favouring_oxic)

@@

## Initial conditions favouring oxic, with noise, no forcing

Figure 2c, d in Bush et al. but with noise.

sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
                          noise_sigma = noise_sigma_0.01)

plot_dynamics(sim_res)
##ggsave("high_CB.png")
```

# Results: Composition/diversity and nature of response

@@

# Results: Varying strength of inhibition

## Introduction

Three parameters set the strength of inhibition (half-inhibition constants):

* h_SR_CB = 300,
* h_O_PB = 100,
* h_O_SB = 100,

These work by the inhibition function:

$\frac{1}{1 + \frac{X}{h_X}}$

```{r}
inhibition
```

Plot some examples:

```{r}
default_parameter_values["h_SR_CB"]
SR <- 10^seq(1.4, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = SR, y = inhibition(SR, default_parameter_values["h_SR_CB"]))) +
  ylab("Multiplicative effect\non CB growth rate")
```

```{r}
default_parameter_values["h_O_PB"]
O <- 10^seq(-0.2, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = O, y = inhibition(O, default_parameter_values["h_O_PB"]))) +
  ylab("Multiplicative effect\non PB growth rate")
```

```{r}
default_parameter_values["h_O_SB"]
O <- 10^seq(-0.2, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = O, y = inhibition(O, default_parameter_values["h_O_SB"]))) +
  ylab("Multiplicative effect\non SB growth rate")
```

```{r}
proportion_change <- 2
default_parameter_values["h_O_SB"]
O <- 10^seq(-0.2, 2.5, length = 100)
ggplot() +
  geom_line(aes(x = O, y = inhibition(O, proportion_change*default_parameter_values["h_O_SB"]))) +
  ylab("Multiplicative effect\non SB growth rate")
```

## Experiment

Increasing the inhibition half saturation constants weakens the inhibition.

```{r}
ssfind_minimum_abundances <- default_minimum_abundances
ssfind_simulation_duration <- 10000
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
ssfind_parameters <- default_parameter_values
ssfind_init_state <- c(N_CB = NA, #this one is varied
           N_PB = NA,
           N_SB = NA,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)

grid_num_mult <- 20
grid_num_N <- 2
grid_num_a <- 100

a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
inhib_const_mults <- 2^seq(-1, 0.4, length=grid_num_mult)

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 1e5
initial_SBs <- 1e5

expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    inhib_const_mult = inhib_const_mults,
                    a_O = a_Os)
```

```{r eval = run_sims}
ss_h1 <- ss_by_a_h_N(expt)

saveRDS(ss_h1, "data/ss_h1.RDS")
```

```{r}
ss_h1 <- readRDS("data/ss_h1.RDS")
```

## Results

### Close to default inhibition

```{r}
inhib_const_mults
focal_inhibition_multiplier <- inhib_const_mults[14]

ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### Double strength inhibition

```{r}
## YES! 0.5 does mean twice as strong inhibition
focal_inhibition_multiplier <- 0.5

ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### Weaker inhibition

```{r}
## YES! 2 does mean half as strong inhibition
focal_inhibition_multiplier <- 2^0.4

ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_h1 %>%
  filter(inhib_const_mult == focal_inhibition_multiplier) %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
x_long <- ss_h1 %>%
  mutate(direction = ifelse(initial_N_CB==1, "up", "down")) %>%
  gather(key = Species, value = Quantity, 3:9) 
x_wide <- x_long %>%
  select(-initial_N_CB) %>%
  spread(key = direction, value=Quantity, drop=T)

resp_summary <- x_wide %>%
  group_by(inhib_const_mult, Species) %>%
  summarise(hyst_1 = mean(abs(log10(up) - log10(down))),
            hyst_2 = get_hysteresis_range(up, down, a),
            nl_up = get_nonlinearity(a, log10(up)),
            nl_down = get_nonlinearity(a, log10(down))) %>%
  ungroup() %>%
  gather(key = resp_measure, value = value, 3:6) %>%
  mutate(
         var_type = ifelse(str_sub(Species, 1, 1)=="N", "Organism", "Substrate")
         )
```

```{r}
resp_summ1 <- resp_summary %>%
  filter(resp_measure %in% c("nl_down", "nl_up")) %>%
  group_by(inhib_const_mult, Species, var_type) %>%
  summarise(ave_nl = mean(value))

resp_summ1 %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_nl)) +
  facet_wrap( ~ Species)

```{r}
focus_on <- "Substrate"
ave_nl_substrate <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(inhib_const_mult) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_substrate %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

```{r}
focus_on <- "Organism"
ave_nl_organism <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(inhib_const_mult) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_organism %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

```{r}
focus_on <- "System"
ave_nl_system <- resp_summ1 %>%
  group_by(inhib_const_mult) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_system %>%
  ggplot() +
  geom_point(aes(inhib_const_mult, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
ggsave("figures/system_nonlinearity.png", width = 3, height = 2)

```{r}
focus_on <- "Substrate"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(inhib_const_mult, resp_measure, var_type) %>%
  summarise(ave_hyst = mean(value))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = inhib_const_mult, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

```{r}
focus_on <- "Organism"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(inhib_const_mult,  resp_measure, var_type) %>%
  summarise(ave_hyst = mean(value))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = inhib_const_mult, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

```{r}
resp_summ4 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2")) %>%
  group_by(resp_measure, inhib_const_mult) %>%
  summarise(mean_hyst = mean(value))

resp_summ4 %>%
  ggplot() +
  geom_point(aes(x = inhib_const_mult, y = mean_hyst)) +
  facet_wrap( ~ resp_measure) +
  ggtitle("Hysteresis in system response.")

```{r}
resp_summ5 <- resp_summ4 %>%
  group_by(inhib_const_mult) %>%
  summarise(mean_mean_hyst = mean(mean_hyst))

resp_summ5 %>%
  ggplot() +
  geom_point(aes(x = inhib_const_mult, y = mean_mean_hyst)) +
  xlab("Organismal tolerance") +
  ylab("Amount of hysteresis") +
  theme_light() +
  ylim(0,2)
ggsave("figures/inhibition_system_hysteresis.png", width = 2.5, height = 2)

# Results: Temporal state switching

## Switch anoxic-oxic-anoxic

## Switch oxic-anoxic-oxic

# Appendices

Rainer: don't change what comes below

## runsteady function not finding correct values

```{r eval = FALSE}

init_state_anoxic_2 <- c(N_CB = 1e1,
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 300,
           SR = 300,
           O = 1e-5,
           P = 1e1)

RS <- runsteady(y = init_state_anoxic_2,
                fun = bushplus_dynamic_model,
                parms = default_parameter_values,
                times = c(0, 1e5),
                stol=1)
RS$y
RS <- runsteady(y = init_state_favouring_oxic,
                fun = bushplus_dynamic_model,
                parms = default_parameter_values,
                times = c(0, 1e5),
                stol=1)
RS$y

## Separatrix

Note that this section not working / updated since the code was made into functions

### Effect of N_CB and oxygen diffusability

```{r eval = FALSE}

noise_sigma <- 0.0

get_final_states_N_CB <- function(x) {

  #give N_CB, N_SB, N_PB and a_O values and put these into state and parameters
  state["N_CB"] <- x["N_CB"]
  parameters["a_O"] <- x["a_O"]

  ## update forcing function
  a_forcing1 <- matrix(ncol=2, byrow=T, data=c(0,
                                             log10(parameters["a_O"]),
                                             max(times),
                                             log10(parameters["a_O"])))
  log10a_forcing2 <- approxfun(x = a_forcing1[,1],
                             y = a_forcing1[,2],
                             method = "linear",
                             rule = 2)

  ## assign the changed a_forcing2 into the global environment, from where the model gets a_forcing2
  assign("log10a_forcing2", log10a_forcing2, envir = .GlobalEnv)

  as.data.frame(ode(y = state,
                    times = times,
                    func = micro_mod_var,
                    parms = parameters,
                    method = "radau",
                    events = list(func=noise_events, time=times)))[2,-1] 
  #gives you values for the last time step
}

times <- seq(0, 50000, by = 50000)

state <- c(N_CB = NA, #this one is varied
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)

#different combinations of N_CB and oxygen diffusability
grid_num <- 50
expt <- expand.grid(N_CB = 10^seq(0, 8, length=grid_num),
                    a_O = 10^seq(-3.5, -2, length=grid_num))

result <- apply(expt, 1, function(x) get_final_states_N_CB(x)) %>%
                          tibble() %>%
                          unnest() %>%
                          mutate(initial_N_CB=expt$N_CB,
                                 a_O=expt$a_O,
                                 highlow=ifelse(N_CB>2e8, 'high', 'low'))

result <- result %>%
  mutate(N_CB = ifelse(N_CB < minimum_abundance, minimum_abundance, N_CB))

result$highlow <- as.factor(result$highlow)
saveRDS(result, "data/CB_oxdiff_hyst.RDS")
```

```{r eval = FALSE}
result <- readRDS("data/CB_oxdiff_hyst.RDS")
```

```{r eval = FALSE}

#plot N_CB / P_CB against a_0
p <- ggplot(result) + 
  geom_line(aes(x=log10(a_O), y=log10(N_CB), color=highlow), size=1.25) +
  ggtitle("Stable states of the system")+
  xlab('log[oxygen diffusivity]') +
  ylab('Initial cyanobacteria abundance')
p <- p + geom_point(data=result, aes(x=log10(a_O), y=log10(initial_N_CB), color=highlow)) + labs(color = "Final CB concentration")
plot(p)
#ggsave("initial conditions_1.png")
```

```{r eval = FALSE}
#show hysterisis effects for all the parameters
#plot oxygen against diffusability
ggplot(result) + 
  geom_point(aes(x=log10(a_O), y=(log10(O))))+
  xlab('log[oxygen diffusivity]') +
  ylab('log[oxygen concentration]')+
  geom_segment(aes(x=log10(result$a_O[839]), y=log10(result$O[839]), xend=log10(result$a_O[838]), yend=log10(result$O[838])), arrow=arrow(), colour="blue", size=0.5)+
  geom_segment(aes(x=log10(result$a_O[1761]), y=log10(result$O[1761]), xend=log10(result$a_O[1762]), yend=log10(result$O[1762])), arrow=arrow(), colour="blue", size=0.5)+
  geom_segment(aes(x=-4.25, y=-0.45, xend=-3.75, yend=0.1), arrow=arrow(), colour="blue", size=0.5)+
  geom_curve(aes(x = -3.75, y = 1.9, xend = -4.25, yend = 1.5), arrow=arrow(), colour='blue', size = 0.5, curvature = -0.2)
#ggsave("hysterisis_CB_1.png")
```

```{r eval = FALSE}
ggplot(result) + 
  geom_point(aes(x=log10(a_O), y=(log10(P))))+
  ggtitle("Hysterisis effect of phoshate concentration")+
  xlab('log[oxygen diffusivity]') +
  ylab('log[phosphate concentration]')
#ggsave("hysterisis_CB_2.png")
```

```{r eval = FALSE}
ggplot(result) + 
  geom_point(aes(x=log10(a_O), y=(log10(SO))))+
  ggtitle("Hysterisis effect of oxidized sulfur concentration")+
  xlab('log[oxygen diffusivity]') +
  ylab('log[oxydized sulfur concentration]')
#ggsave("hysterisis_CB_3.png")
```

```{r eval = FALSE}
ggplot(result) + 
  geom_point(aes(x=log10(a_O), y=(log10(SR))))+
  ggtitle("Hysterisis effect of reduced sulfur concentration")+
  xlab('log[oxygen diffusivity]') +
  ylab('log[reduced sulfur] concentration]')
#ggsave("hysterisis_CB_4.png")
```

These graphs show an at least qualitative match to the result in the paper.

### Effect of PB/SB depending on oxygen diffusability

```{r eval = FALSE}
noise_sigma <- 0.00
get_final_states_N_PB <- function(x) { 
  #give in N_CB, N_SB, N_PB and a_O values and put these into state and parameters
  state["N_PB"] <- x["var"]
  state["N_SB"] <- x["var"]
  parameters["a_O"] <- x["a_O"]

  ## update forcing function
  a_forcing1 <- matrix(ncol=2, byrow=T, data=c(0,
                                             log10(parameters["a_O"]),
                                             max(times),
                                             log10(parameters["a_O"])))
  log10a_forcing2 <- approxfun(x = a_forcing1[,1],
                             y = a_forcing1[,2],
                             method = "linear",
                             rule = 2)
  ## assign the changed a_forcing2 into the global environment, from where the model gets a_forcing2
  assign("log10a_forcing2", log10a_forcing2, envir = .GlobalEnv)

  as.data.frame(ode(y = state,
                         times = times,
                         func = micro_mod_var,
                         parms = parameters,
                         method = "radau", events = list(func=noise_events, time=times)))[2,-1] 
  #gives you values for the last time step
}

times <- seq(0, 50000, by = 5000)
state <- c(N_CB = 5, 
           N_PB = NA,#this one is varied
           N_SB = NA,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)

#different combinations of N_PB and oxygen diffusivity
grid_num <- 50
expt <- expand.grid(var=10^seq(0, 7, length=grid_num),
                    a_O = 10^seq(-3, -2, length=grid_num))

result2 <- apply(expt, 1, function(x) get_final_states_N_PB(x)) %>%
                          tibble() %>%
                          unnest() %>%
                          mutate(initial_N_PB=expt$var,
                                 initial_N_SB=expt$var,
                                 a_O=expt$a_O,
                                 highlow=ifelse(N_PB>1e8, 'high', 'low'))

result2 <- result2 %>%
  mutate(N_PB = ifelse(N_PB < minimum_abundance, minimum_abundance, N_PB))

result2$highlow <- as.factor(result2$highlow)
saveRDS(result2, "data/PBSB_oxdiff_hyst.RDS")
```

```{r eval = FALSE}
result2 <- readRDS("data/PBSB_oxdiff_hyst.RDS")
```

```{r eval = FALSE}
#plot N_PB against a_0
p <- ggplot(result2) + 
  geom_line(aes(x=log10(a_O), y=log10(N_PB), color=highlow), size=1.25) +
  ggtitle("Stable states of the system")+
  xlab('log[oxygen diffusivity]') +
  ylab('Initial PS and SR bacteria concentration')
p <- p + geom_point(data=result2,
                    aes(x=log10(a_O), y=log10(initial_N_PB), color=highlow))
p <- p + labs(color = "Final PB and SB concentration")
plot(p)
#ggsave("initial conditions_2.png")
```

## EWS analyses

Evaluation of the switch: 
Get the functions: 
```{r eval = F}
source('C:/Users/probs/Desktop/Sem 6/Internship/functions_analysis_EWS_2.R')
```

Get state shifts and calculate parameters: 
```{r eval = F}
x3 <- 30 #size of wished rolling window (percentage btw 0 and 100)
x5 <- 50 #size of wished rolling window (percentage btw 0 and 100)

#first switch
name_1='anoxic_oxic 1'
trans1 <- data.frame(out_t %>% filter(time>87500 & time<92500)) #for analysis
trans1_f <- data.frame(out_f %>% filter(time>87500 & time<92500)) #for plotting
n_13 <- round(length(trans1$time)*x3/100) #size of rolling window
t_13 <- length(trans1$time) - n_13 + 1 #number of rolling windows
n_15 <- round(length(trans1$time)*x5/100) #size of rolling window
t_15 <- length(trans1$time) - n_15 + 1 #number of rolling windows
out_df_1_N_CB <- data.frame(trans1$N_CB)
colnames(out_df_1_N_CB) <- c("N_CB")
out_df_1_N_PB <- data.frame(trans1$N_PB)
colnames(out_df_1_N_PB) <- c("N_PB")
out_df_1_N_SB <- data.frame(trans1$N_SB)
colnames(out_df_1_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans1_f) +
  ggtitle("Anoxic_oxic 1")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("anoxic_oxic 1.png")

#second switch
name_2='oxic_anoxic 1'
trans2 <- data.frame(out_t %>% filter(time>335000 & time<360000))
trans2_f <- data.frame(out_f %>% filter(time>335000 & time<360000))
n_23 <- round(length(trans2$time)*x3/100) 
t_23 <- length(trans2$time) - n_23 + 1
n_25 <- round(length(trans2$time)*x5/100) 
t_25 <- length(trans2$time) - n_25 + 1
out_df_2_N_CB <- data.frame(trans2$N_CB)
colnames(out_df_2_N_CB) <- c("N_CB")
out_df_2_N_PB <- data.frame(trans2$N_PB)
colnames(out_df_2_N_PB) <- c("N_PB")
out_df_2_N_SB <- data.frame(trans2$N_SB)
colnames(out_df_2_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans2_f) +
  ggtitle("Oxic_anoxic 1")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("oxic_anoxic 1.png")

#third switch
name_3='oxic_anoxic 2'
trans3 <- data.frame(out2_t %>% filter(time>165000 & time<185000))
trans3_f <- data.frame(out2_f %>% filter(time>165000 & time<185000))
n_33 <- round(length(trans3$time)*x3/100) 
t_33 <- length(trans3$time) - n_33 + 1
n_35 <- round(length(trans3$time)*x5/100) 
t_35 <- length(trans3$time) - n_35 + 1
out2_df_1_N_CB <- data.frame(trans3$N_CB)
colnames(out2_df_1_N_CB) <- c("N_CB")
out2_df_1_N_SB <- data.frame(trans3$N_SB)
colnames(out2_df_1_N_SB) <- c("N_SB")
out2_df_1_N_PB <- data.frame(trans3$N_PB)
colnames(out2_df_1_N_PB) <- c("N_PB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans3_f) +
  ggtitle("Oxic_anoxic 2")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("oxic_anoxic 2.png")

#fourth switch
name_4='anoxic_oxic 2'
trans4 <- data.frame(out2_t %>% filter(time>340000 & time<348500))
trans4_f <- data.frame(out2_f %>% filter(time>340000 & time<348500))
n_43 <- round(length(trans4$time)*x3/100) 
t_43 <- length(trans4$time) - n_43 + 1
n_45 <- round(length(trans4$time)*x5/100) 
t_45 <- length(trans4$time) - n_45 + 1
out2_df_2_N_CB <- data.frame(trans4$N_CB)
colnames(out2_df_2_N_CB) <- c("N_CB")
out2_df_2_N_PB <- data.frame(trans4$N_PB)
colnames(out2_df_2_N_PB) <- c("N_PB")
out2_df_2_N_SB <- data.frame(trans4$N_SB)
colnames(out2_df_2_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans4_f) +
  ggtitle("Anoxic_oxic 2")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("anoxic_oxic 2.png")
```

Rolling window analysis
```{r eval = F}
#analysis for 1st
rollingwindow_analysis(trans1, 2, n_15, t_15, name_1)
rollingwindow_analysis(trans1, 3, n_15, t_15, name_1)
rollingwindow_analysis(trans1, 4, n_15, t_15, name_1)

#analysis for 2nd
rollingwindow_analysis(trans2, 2, n_25, t_25, name_2)
rollingwindow_analysis(trans2, 3, n_25, t_25, name_2)
rollingwindow_analysis(trans2, 4, n_25, t_25, name_2)

#analysis for 3rd
rollingwindow_analysis(trans3, 2, n_35, t_35, name_3)
rollingwindow_analysis(trans3, 3, n_35, t_35, name_3)
rollingwindow_analysis(trans3, 4, n_35, t_35, name_3)

#analysis for 4th
rollingwindow_analysis(trans4, 2, n_45, t_45, name_4)
rollingwindow_analysis(trans4, 3, n_45, t_45, name_4)
rollingwindow_analysis(trans4, 4, n_45, t_45, name_4)

Metric based indicators from the earlywarnings package
```{r eval = F}
#analysis for 1st
earlywarnings_plot(out_df_1_N_CB, name_1, x3)
earlywarnings_plot(out_df_1_N_PB, name_1, x3)
earlywarnings_plot(out_df_1_N_SB, name_1, x3)

#analysis for 2nd
earlywarnings_plot(out_df_2_N_CB, name_2, x3)
earlywarnings_plot(out_df_2_N_PB, name_2, x3)
earlywarnings_plot(out_df_2_N_SB, name_2, x3)

#analysis for 3rd
earlywarnings_plot(out2_df_1_N_CB, name_3, x3)
earlywarnings_plot(out2_df_1_N_PB, name_3, x3)
earlywarnings_plot(out2_df_1_N_SB, name_3, x3)

#analysis for 4th
earlywarnings_plot(out2_df_2_N_CB, name_4, x3)
earlywarnings_plot(out2_df_2_N_PB, name_4, x3)
earlywarnings_plot(out2_df_2_N_SB, name_4, x3)
```

Conditional Heteroskedasticity 
```{r eval = F}
#analysis for 1st
ch <- condhet_analysis(out_df_1_N_CB, name_1)
ch <- condhet_analysis(out_df_1_N_PB, name_1)
ch <- condhet_analysis(out_df_1_N_SB, name_1)

>>>>

---
title: "Partial reproduction of Bush et al (2017) Oxic-anoxic regime shifts mediated by feedbacks between biogeochemical processes and microbial community dynamics"
author: "Owen Petchey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducing some results of Bush et al 2017}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

@@

library(tidyverse)
library(deSolve)
library(rootSolve)
library(here)
library(microxanox)

## Set random number generator seed
## Is redundant if there are no stochastic components in the model
set.seed(13)

run_steadystate_sims <- FALSE
```

@@

Only one set of parameter values is used throughout. It gives the possibility for alternate stable states.

@@

  ## oxygen production per microbial cell
  Pr_CB = 6e-9,

@@

Specify the default set of initial conditions (here favouring anoxic organisms).

Default to the following duration and interval between samples:

Specified by giving a vector of values of $a_0$. These are then automatically spaced equally along the duration of the simulation. Here we specify no forcing of $a_0$.

@@

Default is to ensure abundances are not below minimum, and to add the amount of noise specified below (none in this vignette).

These default are in the definition function `default_event_definition()`

@@

Here set to an event every 10 time steps.

@@

Default to no noise added to substrate concentrations:

@@

Default to following minimum abundances:

@@

@@

Remove this section when the rest definitely works.

```{r eval = FALSE}
dynamic_model = bushplus_dynamic_model
event_definition = default_event_definition
parameter_values = default_parameter_values
event_interval = default_event_interval
noise_sigma = default_noise_sigma
minimum_abundances = default_minimum_abundances
sim_duration = default_sim_duration
sim_sample_interval = default_sim_sample_interval
log10a_series = default_log10a_series
initial_state = default_initial_state
solver_method = "radau"

times <- seq(0,
             sim_duration,
             by = sim_sample_interval)

event_times <- seq(min(times),
                   max(times),
                   by = event_interval)

log10a_forcing <- matrix(ncol=2,
                         byrow=F,
                         data=c(ceiling(seq(0,
                                            max(times),
                                            length=length(log10a_series))),
                                log10a_series))

log10a_forcing <- approxfun(x = log10a_forcing[,1],
                            y = log10a_forcing[,2],
                            method = "linear",
                            rule = 2)
assign("log10a_forcing_func", log10a_forcing, envir = .GlobalEnv)

out <- as.data.frame(ode(y = initial_state,
                         times = times,
                         func = dynamic_model,
                         parms = parameter_values,
                         method = solver_method,
                         events = list(func=event_definition,
                                       time=event_times)))

bushplus_dynamic_model(1, initial_state, default_parameter_values)

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
                           event_definition = default_event_definition,
                           parameter_values = default_parameter_values,
                           event_interval = default_event_interval,
                           noise_sigma = default_noise_sigma,
                           minimum_abundances = default_minimum_abundances,
                           sim_duration = default_sim_duration,
                           sim_sample_interval = default_sim_sample_interval,
                           log10a_series = default_log10a_series,
                           initial_state = default_initial_state,
                           solver_method = "radau")

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model)

plot_dynamics(sim_res)

@@

# Results

@@

parameters <- default_parameter_values
state <- init_state_favouring_anoxic
bushplus_dynamic_model(100, state, parameters)

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
  initial_state = init_state_favouring_anoxic)
plot_dynamics(simulation_result = sim_res)

@@

@@

sim_res <- run_simulation(dynamic_model = bushplus_dynamic_model,
  initial_state = init_state_favouring_oxic)

@@

## Stable states

```{r eval = run_steadystate_sims}
ssfind_minimum_abundances <- rep(0, 3)
names(ssfind_minimum_abundances) <- c("CB", "PB", "SB") 
ssfind_simulation_duration <- 10000
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
ssfind_parameters <- default_parameter_values
# ssfind_init_state <- c(N_CB = NA, 
#                        N_PB = NA,
#                        N_SB = NA,
#                        SO = 500,
#                        SR = 50,
#                        O = 30,
#                        P = 10)
##ssfind_init_state <- default_initial_state
ssfind_init_state <- init_state_favouring_anoxic
ssfind_init_state <- c(N_CB = NA, #this one is varied
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)
grid_num_a <- 2
a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
grid_num_N <- 2

ss_CB_SB_PB <- ss_by_a_N(expt)
save(ss_CB_SB_PB, file = here("vignettes/data/ss_CB_SB_PB.Rda"))
```

```{r}
load(file = here("vignettes/data/ss_CB_SB_PB.Rda"))
ss_CB_SB_PB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species),
            size=0.8) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  theme_light() +
  ylim(0,10) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="CB, SB, & PB")  %>%
  print()
##ggsave("figures/CB_SB_PB.png", width = 3, height = 2)

```{r}
p1 <- ss_CB_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
print(p1)
```

## Temporal state switching

### Anoxic-oxic-anoxic

### Oxic-anoxic-oxic

## Composition and nature of response

@@
